---
title: "Tobacco consumption prediction for 2021"
author: "Elías Garza, Diego Rodríguez, Federico Medina"
date: "28/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning=FALSE,message=FALSE}

library(dplyr)
library(fpp2)
library(readr)
library(ggplot2)
library(forecast)
library(forecastHybrid)
library(gbm)
library(nnfor)
```

\section{Reading Data}

First of all, we need to read the data and analyze it to understand the structure of the dataset and decide what to do. Here, we realized that the dataset is conformed by the data of 13 tobacco products in 20 years. Also, we decided to create a new column named 'item' which concatenate other 2 variables (Submeasure and 'Data Value Unit'), in this way is easier to work with each product given that some of them have the same submeasure.

```{r, warning=FALSE,message=FALSE}
Tmatrix <- read_csv("Tobacco_Consumption.csv")
Tdata<-as.data.frame(Tmatrix)
Tdata$item<-paste(Tdata$Submeasure," in ",Tdata$`Data Value Unit`)
```

\section{Dividing by products}

Next step, we divided the dataset in 13 different dataframes for each product using the new variable 'item' so we can work with them separately.

```{r}
Products<-list()
for (i in Tdata$item[1:13]){
  Products<-c(Products,list(filter(Tdata, item==i)))
}
names(Products)<-Tdata$item[1:13]
```

and created df of Totals, Imports and Domestic per Capita per Product since these are de variables with which we will work. For this, we created new vectors for the per capita values, since the ones on the original dataframe appear to be rounded and some of this values cause problems, more than anything the 0s.

```{r}
totalsPerCapita<-Products[[1]]%>%select(11)/Products[[1]]$Population
importsPerCapita<-Products[[1]]%>%select(10)/Products[[1]]$Population
domesticPerCapita<-Products[[1]]%>%select(9)/Products[[1]]$Population
for(j in c(2:13)){
  totalsPerCapita<-cbind(totalsPerCapita,Products[[j]]%>%select(11)/Products[[j]]$Population)
  importsPerCapita<-cbind(importsPerCapita,Products[[j]]%>%select(10)/Products[[j]]$Population)
  domesticPerCapita<-cbind(domesticPerCapita,Products[[j]]%>%select(9)/Products[[j]]$Population)
}
```

From the next plots we can extract some information:

\begin{itemize}
  \item Something happened in 2008 that affected tobacco consumption in most of its forms.
  \item From snuff and chewing tobacco (the only 2 forms of noncombustible tobacco), it seems like chewing tobacco is loosing popularity while snuff gaining.
  \item Both cigars and large cigars are preferable from imports to domestic ones. This is important to point out since for every other product there is a big preference for domestic production.

```{r}
for(i in c(1:13)){
  print(
      ggplot(data=Products[[i]], aes(x=c(2000:2020))) +
      geom_line(aes(y =totalsPerCapita[[i]],color='Total'))+
      geom_line(aes(y=importsPerCapita[[i]],color='Imports'))+
      geom_line(aes(y=domesticPerCapita[[i]],color='Domestic'))+
      xlab('Years')+ylab('Totals Per Capita')+
      labs(title=names(Products)[i])+
      scale_color_manual(name='Legend',values = c('Total' = "black", "Imports" = "blue",'Domestic'='red')))
}
```

\section{Generating Training and Testing data}

To start with the prediction section, first its needed to create a training and testing data. We'll use data from 2000 to 2016 as training and from 2017 to 2020 for testing.

```{r}
trainTotals<-list()
testTotals<-list()
for(i in c(1:13)){
  trainTotals<-c(trainTotals, list(ts(head(totalsPerCapita[[i]],17),start=c(2000),end=c(2016),frequency = 1)))
  testTotals<-c(testTotals, list(ts(tail(totalsPerCapita[[i]],4),start=c(2017),end=c(2020),frequency = 1)))
}

```

Function to get the Mean Squared Error from 2 vectors
```{r}
MSE<- function (v1,v2){
  return(sum((v1-v2)^2)/length(v1))
}
```

Now everything is ready to start with the models. 

\section{Auto-ARIMA}
```{r}
arimaErrors<-c()
for(i in c(1:13)){
  #Training and making forecast until 2021 using AUTO-Arima 
  sarima_ts<-auto.arima(trainTotals[[i]])
  arima_model<-forecast::forecast(sarima_ts,h=5)
  
  #Plotting prediction and testing data (red for testing data)
  plot(arima_model)
  lines(testTotals[[i]],col='red')
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-arima_model$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  #Saving MSE in arimaError vector
  arimaErrors<-c(arimaErrors,MSE(prediction,test))
}
```

\section{Neural Network Autoregression}
```{r}
nnErrors<-c()
for(i in c(1:13)){
  #Training model
  fit<-nnetar(trainTotals[[i]],lambda='auto')
  nn_model<-forecast::forecast(fit,h=5)
  
  #Plotting prediction and testing data (red for testing data)
  plot(nn_model)
  lines(testTotals[[i]],col='red')  
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-nn_model$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  #Saving MSE in nnError vector
  nnErrors<-c(nnErrors,MSE(prediction,test))
}
```

\section{Hybrid Model}
```{r}
hybErrors<-c()
for(i in c(1:13)){
  #Training and making forecast
  hyb_mod<- hybridModel(trainTotals[[i]])
  hyb_forecast <- forecast::forecast(hyb_mod,5)
  
  #Plotting prediction and testing data (red for testing data)
  plot(hyb_forecast)  
  lines(testTotals[[i]],col='red') 
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-hyb_forecast$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  #Saving MSE in hybError vector
  hybErrors<-c(hybErrors,MSE(prediction,test))
}
```

\section{Multilayer Perceptron Model}
```{r}
mlpErrors<-c()
for(i in c(1:13)){
  #Training
  mlp_fit<-mlp(trainTotals[[i]])
  mlp_model<-forecast::forecast(mlp_fit,5)
  
  #Plotting
  plot(mlp_model)
  lines(testTotals[[i]],col='red')
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-mlp_model$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  ##Saving MSE in mlpError vector
  mlpErrors<-c(mlpErrors,MSE(prediction,test))
}
```

\section{Exponential Smoothing}
```{r}
library(smooth)
library(greybox)

expErrors<-c()
for (i in c(1:13)){
  #Generating adn plotting model
  exp_model<-es(trainTotals[[i]], h=5, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  ##Saving MSE in expError vector
  expErrors<-c(expErrors,MSE(prediction,test))
}
```

\section{Simple Moving Average}
```{r}
smaErrors<-c()
for (i in c(1:13)){
  #Generating and plotting model
  sma_model<-sma(trainTotals[[i]], h=5, order=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-sma_model$forecast%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  ##Saving MSE in smaError vector
  smaErrors<-c(smaErrors,MSE(prediction,test))
}
```

\section{Polinomial Regression}
```{r}
polErrors<-c()
for(h in c(1:13)){
  #Load and plot the data
  polydf <- data.frame(year=c(2000:2016),value=trainTotals[[h]]%>%as.numeric())
  
  #randomly shuffle data
  polydf.shuffled <- polydf[sample(nrow(polydf)),]
  
  #define number of folds to use for k-fold cross-validation
  K <- 10 
  
  #define degree of polynomials to fit
  degree <- 5
  
  #create k equal-sized folds
  folds <- cut(seq(1, nrow(polydf.shuffled)) , breaks=K , labels=FALSE)
  
  #create object to hold MSE's of models
  mse = matrix(data=NA,nrow=K,ncol=degree)
  
  #Perform K-fold cross validation
  for(i in 1:K){
      
      #define training and testing data
      testData <- data.frame(year=c(2017:2020),value=testTotals[[h]]%>%as.numeric())
      trainData <- data.frame(year=c(2000:2016),value=trainTotals[[h]]%>%as.numeric())
      
      #use k-fold cv to evaluate models
      for (j in 1:degree){
          fit.train = lm(value ~ poly(year,j), data=trainData)
          fit.test = predict(fit.train, newdata=testData)
          mse[i,j] = mean((fit.test-testData$value)^2) 
      }
  }
  
  #find MSE for each degree 
  mmse =colMeans(mse)
  #determine which is the better degree
  mdegree = which.min(mmse)
  
  # Make predictions
  model <- lm(value ~ poly(year, mdegree), data = polydf)
  predictions <- model %>% predict(data.frame('year'=c(2017:2021)))
  predictionsdf <- data.frame('year' = c(2017: 2021), 'value' = predictions)
  totaldf <- rbind(polydf, predictionsdf )
  
  print(ggplot(totaldf, aes(x=year, y=value)) + 
            geom_point() +
            stat_smooth(method='lm', formula = y ~ poly(x,mdegree), size = 1)+
            xlab('year') +
            ylab('value'))
  
  #Saving MSE in vector
  polErrors<-c(polErrors,MSE(predictions[1:4],testTotals[[i]]%>%as.numeric()))
}

```


\section{Error dtaframe generation}

Once models are done, we have to measure their error and compare them with each other.

```{r}
#Generating dataframe of all MSEs
Error<-cbind(arimaErrors,nnErrors,hybErrors,mlpErrors,expErrors,smaErrors,polErrors)%>%t()%>%data.frame()
names(Error)<-Products%>%names()
Error
```

Getting best model for every type of tobacco

```{r}
for(i in c(1:13)){
  print(rownames(Error)[which.min(Error[,i])])
}
```

2021 Final predictions per capita with 10% training set

```{r}
#Generating new training set
trainTotals10<-list()
testTotals10<-list()
for(i in c(1:13)){
  trainTotals10<-c(trainTotals10, list(ts(head(totalsPerCapita[[i]],19),start=c(2000),end=c(2018),frequency = 1)))
  testTotals10<-c(testTotals10, list(ts(tail(totalsPerCapita[[i]],2),start=c(2019),end=c(2020),frequency = 1)))
}

#Creating array to store the 2021 forecasts and errors
forecast2021<-c()
predictionMSE<-c()
```

\section{Chewing Tobacco in Pounds forecast}
```{r}
 #Generating adn plotting model
  exp_model<-es(trainTotals10[[1]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[1]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Cigarrates Removal in Cigarrates forecast}
```{r}
 #Generating adn plotting model
  exp_model<-es(trainTotals10[[2]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[2]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Total Cigars in Cigars Forecast}
```{r}
  #Generating and plotting model
  sma_model<-sma(trainTotals10[[3]], h=3, order=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-sma_model$forecast%>%as.numeric()%>%tail(3)%>%head(1)
  test<-testTotals10[[3]]%>%as.numeric()
  
  #Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,sma_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Total loose Tobacco in Cigarrates equivalent forecast}
The data wont change because we move between units so if we do the same with Total loose Tobacco in Pounds it will show the same graph and prediction but scaled. 
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[4]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[4]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Small Cigars in Cigars forecast}
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[6]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[6]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Pippe Tobacco in Pounds forecast}
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[7]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[7]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Roll-your-own Tobacco in cigarrate equivalents forecast}
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[8]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[8]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Large Cigars in Cigars forecast}
```{r}
  library(forecast)
  library(forecastHybrid)
  library(fpp2)
  library(nnfor)
  #Training model
  fit<-nnetar(trainTotals10[[10]],lambda='auto')
  nn_model<-forecast::forecast(fit,h=3)
  
  #Plotting prediction and testing data (red for testing data)
  plot(nn_model)
  lines(testTotals10[[10]],col='red')  
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-nn_model$fitted%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[10]]%>%as.numeric()
  
  #Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,nn_model$fittedt%>%as.numeric()%>%tail(1))
```

\section{Snuff in Pounds forecast}
```{r}
  #Training
  mlp_fit<-mlp(trainTotals10[[12]])
  mlp_model<-forecast::forecast(mlp_fit,3)
  
  #Plotting
  plot(mlp_model)
  lines(testTotals10[[12]],col='red')
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-mlp_model$fitted%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[12]]%>%as.numeric()
  
  #Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Pipe Tobacco in cigarratte equivalents forecast}
```{r}
  library(smooth)
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[11]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[11]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

\section{Totals by model}
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[13]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[13]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```



