---
title: "test"
author: "eliasgarzav"
date: "28/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(prophet)
library(dplyr)
library(fpp2)
library(readr)
library(ggplot2)
library(forecast)
library(forecastHybrid)
library(gbm)
library(nnfor)
```

Reading Data
```{r}
Tmatrix <- read_csv("Tobacco_Consumption.csv")
Tdata<-as.data.frame(Tmatrix)
Tdata$item<-paste(Tdata$Submeasure," in ",Tdata$`Data Value Unit`)
```

Dividing by products
```{r}
Products<-list()
for (i in Tdata$item[1:13]){
  Products<-c(Products,list(filter(Tdata, item==i)))
}
names(Products)<-Tdata$item[1:13]
```

Create df of Totals, Imports and Domestic per Capita per Product
```{r}
totalsPerCapita<-Products[[1]]%>%select(11)/Products[[1]]$Population
importsPerCapita<-Products[[1]]%>%select(10)/Products[[1]]$Population
domesticPerCapita<-Products[[1]]%>%select(9)/Products[[1]]$Population
for(j in c(2:13)){
  totalsPerCapita<-cbind(totalsPerCapita,Products[[j]]%>%select(11)/Products[[j]]$Population)
  importsPerCapita<-cbind(importsPerCapita,Products[[j]]%>%select(10)/Products[[j]]$Population)
  domesticPerCapita<-cbind(domesticPerCapita,Products[[j]]%>%select(9)/Products[[j]]$Population)
}
```


```{r}
for(i in c(1:13)){
  print(
      ggplot(data=Products[[i]], aes(x=c(2000:2020))) +
      geom_line(aes(y =totalsPerCapita[[i]],color='Total'))+
      geom_line(aes(y=importsPerCapita[[i]],color='Imports'))+
      geom_line(aes(y=domesticPerCapita[[i]],color='Domestic'))+
      xlab('Years')+ylab('Totals Per Capita')+
      labs(title=names(Products)[i])+
      scale_color_manual(name='Legend',values = c('Total' = "black", "Imports" = "blue",'Domestic'='red')))
}
```

Generating Training and Testing data
```{r}
trainTotals<-list()
testTotals<-list()
for(i in c(1:13)){
  trainTotals<-c(trainTotals, list(ts(head(totalsPerCapita[[i]],17),start=c(2000),end=c(2016),frequency = 1)))
  testTotals<-c(testTotals, list(ts(tail(totalsPerCapita[[i]],4),start=c(2017),end=c(2020),frequency = 1)))
}

```

Function to get the Mean Squared Error from 2 vectors
```{r}
MSE<- function (v1,v2){
  return(sum((v1-v2)^2)/length(v1))
}
```


Auto-ARIMA
```{r}
arimaErrors<-c()
for(i in c(1:13)){
  #Training and making forecast until 2021 using AUTO-Arima 
  sarima_ts<-auto.arima(trainTotals[[i]])
  arima_model<-forecast(sarima_ts,h=5)
  
  #Plotting prediction and testing data (red for testing data)
  plot(arima_model)
  lines(testTotals[[i]],col='red')
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-arima_model$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  #Saving MSE in arimaError vector
  arimaErrors<-c(arimaErrors,MSE(prediction,test))
}
```

Neural Network Autoregression
```{r}
nnErrors<-c()
for(i in c(1:13)){
  #Training model
  fit<-nnetar(trainTotals[[i]],lambda='auto')
  nn_model<-forecast(fit,h=5)
  
  #Plotting prediction and testing data (red for testing data)
  plot(nn_model)
  lines(testTotals[[i]],col='red')  
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-nn_model$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  #Saving MSE in nnError vector
  nnErrors<-c(nnErrors,MSE(prediction,test))
}
```

Hybrid Model
```{r}
hybErrors<-c()
for(i in c(1:13)){
  #Training and making forecast
  hyb_mod<- hybridModel(trainTotals[[i]])
  hyb_forecast <- forecast(hyb_mod,5)
  
  #Plotting prediction and testing data (red for testing data)
  plot(hyb_forecast)  
  lines(testTotals[[i]],col='red') 
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-hyb_forecast$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  #Saving MSE in hybError vector
  hybErrors<-c(hybErrors,MSE(prediction,test))
}
```

Multilayer Perceptron Model
```{r}
mlpErrors<-c()
for(i in c(1:13)){
  #Training
  mlp_fit<-mlp(trainTotals[[i]])
  mlp_model<-forecast(mlp_fit,5)
  
  #Plotting
  plot(mlp_model)
  lines(testTotals[[i]],col='red')
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-mlp_model$fitted%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  ##Saving MSE in mlpError vector
  mlpErrors<-c(mlpErrors,MSE(prediction,test))
}
```

Exponential Smoothing
```{r}
library(smooth)
library(greybox)

expErrors<-c()
for (i in c(1:13)){
  #Generating adn plotting model
  exp_model<-es(trainTotals[[i]], h=5, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  ##Saving MSE in expError vector
  expErrors<-c(expErrors,MSE(prediction,test))
}
```

Simple Moving Average
```{r}
smaErrors<-c()
for (i in c(1:13)){
  #Generating and plotting model
  sma_model<-sma(trainTotals[[i]], h=5, order=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-sma_model$forecast%>%as.numeric()%>%tail(5)%>%head(4)
  test<-testTotals[[i]]%>%as.numeric()
  
  ##Saving MSE in smaError vector
  smaErrors<-c(smaErrors,MSE(prediction,test))
}
```

Polinomial Regression
```{r}
polErrors<-c()
for(h in c(1:13)){
  #Load and plot the data
  polydf <- data.frame(year=c(2000:2016),value=trainTotals[[h]]%>%as.numeric())
  
  #randomly shuffle data
  polydf.shuffled <- polydf[sample(nrow(polydf)),]
  
  #define number of folds to use for k-fold cross-validation
  K <- 10 
  
  #define degree of polynomials to fit
  degree <- 5
  
  #create k equal-sized folds
  folds <- cut(seq(1, nrow(polydf.shuffled)) , breaks=K , labels=FALSE)
  
  #create object to hold MSE's of models
  mse = matrix(data=NA,nrow=K,ncol=degree)
  
  #Perform K-fold cross validation
  for(i in 1:K){
      
      #define training and testing data
      testData <- data.frame(year=c(2017:2020),value=testTotals[[h]]%>%as.numeric())
      trainData <- data.frame(year=c(2000:2016),value=trainTotals[[h]]%>%as.numeric())
      
      #use k-fold cv to evaluate models
      for (j in 1:degree){
          fit.train = lm(value ~ poly(year,j), data=trainData)
          fit.test = predict(fit.train, newdata=testData)
          mse[i,j] = mean((fit.test-testData$value)^2) 
      }
  }
  
  #find MSE for each degree 
  mmse =colMeans(mse)
  #determine which is the better degree
  mdegree = which.min(mmse)
  
  # Make predictions
  model <- lm(value ~ poly(year, mdegree), data = polydf)
  predictions <- model %>% predict(data.frame('year'=c(2017:2021)))
  predictionsdf <- data.frame('year' = c(2017: 2021), 'value' = predictions)
  totaldf <- rbind(polydf, predictionsdf )
  
  print(ggplot(totaldf, aes(x=year, y=value)) + 
            geom_point() +
            stat_smooth(method='lm', formula = y ~ poly(x,mdegree), size = 1)+
            xlab('year') +
            ylab('value'))
  
  #Saving MSE in vector
  polErrors<-c(polErrors,MSE(predictions[1:4],testTotals[[i]]%>%as.numeric()))
}

```

Error dtaframe generation
```{r}
#Generating dataframe of all MSEs
Error<-cbind(arimaErrors,nnErrors,hybErrors,mlpErrors,expErrors,smaErrors,polErrors)%>%t()%>%data.frame()
names(Error)<-Products%>%names()
Error
```

Getting best model for every type of tobacco
```{r}
for(i in c(1:13)){
  print(rownames(Error)[which.min(Error[,i])])
}
```

2021 Final predictions per capita with 10% training set
```{r}
#Generating new training set
trainTotals10<-list()
testTotals10<-list()
for(i in c(1:13)){
  trainTotals10<-c(trainTotals10, list(ts(head(totalsPerCapita[[i]],19),start=c(2000),end=c(2018),frequency = 1)))
  testTotals10<-c(testTotals10, list(ts(tail(totalsPerCapita[[i]],2),start=c(2019),end=c(2020),frequency = 1)))
}

#Creating array to store the 2021 forecasts and errors
forecast2021<-c()
predictionMSE<-c()
```

## Chewing Tobacco in Pounds
```{r}
 #Generating adn plotting model
  exp_model<-es(trainTotals10[[1]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[1]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Cigarrates Removal in Cigarrates
```{r}
 #Generating adn plotting model
  exp_model<-es(trainTotals10[[2]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[2]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Total Cigars in Cigars
```{r}
  #Generating and plotting model
  sma_model<-sma(trainTotals10[[3]], h=3, order=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-sma_model$forecast%>%as.numeric()%>%tail(3)%>%head(1)
  test<-testTotals10[[3]]%>%as.numeric()
  
  #Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,sma_model$forecast%>%as.numeric()%>%tail(1))
```

## Total loose Tobacco in Cigarrates equivalent
The datra wont change because we move between units so if we do the same with Total loose Tobacco in Pounds it will show the same graph and prediction but scaled. 
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[4]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[4]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Small Cigars in Cigars
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[6]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[6]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

##Pippe Tobacco in Pounds
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[7]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[7]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Roll-your-own Tobacco in cigarrate equivalents
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[8]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[8]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Large Cigars in Cigars
```{r}
  library(forecast)
  library(forecastHybrid)
  library(fpp2)
  library(nnfor)
  #Training model
  fit<-nnetar(trainTotals10[[10]],lambda='auto')
  nn_model<-forecast(fit,h=3)
  
  #Plotting prediction and testing data (red for testing data)
  plot(nn_model)
  lines(testTotals10[[10]],col='red')  
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-nn_model$fitted%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[10]]%>%as.numeric()
  
  #Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,nn_model$fittedt%>%as.numeric()%>%tail(1))
```

## Snuff in Pounds
```{r}
  #Training
  mlp_fit<-mlp(trainTotals10[[12]])
  mlp_model<-forecast(mlp_fit,3)
  
  #Plotting
  plot(mlp_model)
  lines(testTotals10[[12]],col='red')
  
  #Getting MSE (the head and tail are used to get from 2017-2020)
  prediction<-mlp_model$fitted%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[12]]%>%as.numeric()
  
  #Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Pipe Tobacco in cigarratte equivalents
```{r}
  library(smooth)
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[11]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[11]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```

## Totals by model
```{r}
  #Generating adn plotting model
  exp_model<-es(trainTotals10[[13]], h=3, holdout=FALSE, interval=TRUE, silent=FALSE)
  
  #Getting MSE (the head and tail are used to get from 2019-2020)
  prediction<-exp_model$forecast%>%as.numeric()%>%tail(3)%>%head(2)
  test<-testTotals10[[13]]%>%as.numeric()
  
  ##Saving MSE and forecast
  predictionMSE<-c(predictionMSE,MSE(prediction,test))
  forecast2021<-c(forecast2021,exp_model$forecast%>%as.numeric()%>%tail(1))
```



